---
title: "PROG8430-23W-Assign03"
author: "Maulik Patel"
date: "2023-02-18"
output: pdf_document
---

Set the work directory. 
This MUST be done for every submission
This will determine the default location of all of your data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#This sets the working directory
knitr::opts_knit$set(root.dir = 'C:/Users/mauli/OneDrive/Documents/BDSA/R/Assignment 3')

```

This section is for the basic set up.
It will clear all the plots, the console and the workspace.
It also sets the overall format for numbers.

```{r}
if(!is.null(dev.list())) dev.off()
cat("\014") 
rm(list=ls())
options(scipen=9)
```
This section loads and attaches all the necessary packages.

```{r}
#Load packages
#For Excel
if(!require(pastecs)){install.packages("pastecs")}
library("pastecs")

if(!require(lattice)){install.packages("lattice")}
library("lattice")

if(!require(readxl)){install.packages("readxl")}
library("readxl")

if(!require(ggplot2)){install.packages("ggplot2")}
library("ggplot2")

if(!require(tinytex)){install.packages("tinytex")}
library("tinytex")
```

1 Data Transformation
  1. Rename all variables with your initials appended (just as was done in
     assignment 1)
    
    ```{r}

#reading the file
fcb_MP <- read.table(file = "PROG8430-23W-Assign03.txt", header=TRUE, sep=",")

#transforming it into a Data frame
fcb_MP <- as.data.frame(fcb_MP)
head(fcb_MP)

#to check the datatype we used class() function
class(fcb_MP)

```

```{r}

#Renaming all variables in table with my name initials
colnames(fcb_MP) <- c("Food_MP", "Enter_MP", "Edu_MP", "Trans_MP", 
                     "Work_MP", "House_MP", "Oth_MP")

#after printing the data frame we can see that the names of the columns are updated with my initials
print(head(fcb_MP))


#using dim() function to get the dimensions of the data frame
dim(fcb_MP)
#we can see that there are 1059 rows and 7 columns in the data frame
```
Descriptive Data Analysis
  1. Create graphical summaries of the data (as demonstrated in class:
     boxplots, histograms or density plots) and comment on any
     observations you make.
```{r}

densityplot(~fcb_MP$Food_MP, xlab = "Food", main="Density plot of of % income spent on Food")
densityplot(~fcb_MP$House_MP, xlab = "House", main="Density plot of % income spent on House")
densityplot(~fcb_MP$Work_MP, xlab = "Work", main="Density plot of % income spent on Work")
densityplot(~fcb_MP$Edu_MP, xlab = "Education", main="Density plot of % income spent on Education")
densityplot(~fcb_MP$Trans_MP, xlab = "Transportation", main="Density plot of % income spent on Transportation")
densityplot(~fcb_MP$Oth_MP, xlab = "Others", main="Density plot of % income spent on Others")
densityplot(~fcb_MP$Enter_MP, xlab = "Others", main="Density plot of % income spent on Entertainment")

```
#densityplots of all the variables depict that the data is not normally distributed but we can see outliers only in two variables from the set of seven variables


Exploring the Data
```{r}

par(mfrow=c(2,2))
for (i in 1:ncol(fcb_MP)) {
  if (is.numeric(fcb_MP[,i])) {
      boxplot(fcb_MP[i], main=names(fcb_MP)[i],
              horizontal=TRUE, pch=10)
  }
}

par(mfrow=c(1,1))

```
2. Standardize all of the variables using either of the two functions
     demonstrated in class. Describe why you chose the method you did.

#we will prefer this method because the data here contains very negligible set outliers which doesn't affect this formula     
```{r}

Normal_MP <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}


fcb_MP$Food_norm_MP <- Normal_MP(fcb_MP$Food_MP)

fcb_MP$Work_norm_MP <- Normal_MP(fcb_MP$Work_MP)

fcb_MP$Enter_norm_MP <- Normal_MP(fcb_MP$Enter_MP)

fcb_MP$Edu_norm_MP <- Normal_MP(fcb_MP$Edu_MP)

fcb_MP$Trans_norm_MP <- Normal_MP(fcb_MP$Trans_MP)

fcb_MP$House_norm_MP <- Normal_MP(fcb_MP$House_MP)

fcb_MP$Oth_norm_MP <- Normal_MP(fcb_MP$Oth_MP)

head(fcb_MP)

```
Clustering
  
  Using the K-Means procedure as demonstrated in class, create
  clusters with k=2,3,4,5,6,7.
  You will be using only two variables as your centroids (House and
  Food)
  1. Create segmentation/cluster schemes for k=2,3,4,5,6,7.
```{r}

#creating clusters 
#initializing using cluster_MP with 7
cluster_MP <- 7
mc_MP <- c(2:cluster_MP)
mss_MP <- rep(0,cluster_MP-1)

```
  2. Create the WSS plots as demonstrated in class and select a
     suitable k value based on the “elbow”. [NOTE – Use the code
     that I provided to do this. Using other functions will yield
     different results.]
     
#creating for all values of k and observing that the values of K changing each time
```{r}

#for k=2
k=2

#we used columns food and house that are 8 and 13
Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_MP

#creating cluster column
fcb_MP$cluster <- factor(Clstr_MP$cluster)
head(fcb_MP)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

```
```{r}

#for k=3
k=3

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_MP

fcb_MP$cluster <- factor(Clstr_MP$cluster)
head(fcb_MP)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

```

```{r}
#for k=4
k=4

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_MP

fcb_MP$cluster <- factor(Clstr_MP$cluster)
head(fcb_MP)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

```
```{r}
#for k=5
k=5

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_MP

fcb_MP$cluster <- factor(Clstr_MP$cluster)   # Adding Cluster tags to variables
head(fcb_MP)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

```

```{r}

#for k=6
k=6

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_MP

fcb_MP$cluster <- factor(Clstr_MP$cluster)
head(fcb_MP)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

```

```{r}
#for k=7
k=7

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_MP

fcb_MP$cluster <- factor(Clstr_MP$cluster)
head(fcb_MP)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

```



```{r}
#creating Elbow chart for clusters above to observe the bend

plot(2:cluster_MP, mss_MP,
     type="b", pch = 19, frame = FALSE,
     main="Elbow Chart for Clusters",
     xlab="Values of K",
     ylab="Total Within-Clusters SS",
     ylim=c(0,max(mss_MP)))


```
Evaluation of Clusters
  1. Based on the “k” chosen above, create a scatter plot showing the
     clusters and colour-coded datapoints for each of “k-1”, “k”, “k+1”. For
     example, if you think the “elbow” is at k=4 create the charts for k=3,
     k=4 and k=5.

#scatterplots for k-1, k, k+1
#we have choosen k=3 from the elbow chart as we can see it bending at k=3
```{r}
#scatter plot for k-1: 3-1=2
k=2

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_MP

fcb_MP$cluster <- factor(Clstr_MP$cluster)   # Adding Cluster tags to variables
head(fcb_MP)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

plot(fcb_MP$Food_norm_MP, fcb_MP$House_norm_MP,
     col=fcb_MP$cluster, pch=as.numeric(fcb_MP$cluster), xlab='Food', ylab='House', main = "Scatterplot for K=2")
points(centers$Food_norm_MP, centers$House_norm_MP,
       col=centers$cluster, pch=as.numeric(centers$cluster),
       cex=5, lwd=3, xlab='Food', ylab='House')
```
```{r}
#scatter plot for k: 3
k=3

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)

fcb_MP$cluster <- factor(Clstr_MP$cluster)   # Adding Cluster tags to variables

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

plot(fcb_MP$Food_norm_MP, fcb_MP$House_norm_MP,
     col=fcb_MP$cluster, pch=as.numeric(fcb_MP$cluster),xlab='Food', ylab='House', main = "Scatterplot for K=3")
points(centers$Food_norm_MP, centers$House_norm_MP,
       col=centers$cluster, pch=as.numeric(centers$cluster),
       cex=5, lwd=3)
```

```{r}

#scatter plot for k+1: 3+1=4
k=4

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)

fcb_MP$cluster <- factor(Clstr_MP$cluster)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

plot(fcb_MP$Food_norm_MP, fcb_MP$House_norm_MP,
     col=fcb_MP$cluster, pch=as.numeric(fcb_MP$cluster),xlab='Food', ylab='House', main = "Scatterplot for K=4")
points(centers$Food_norm_MP, centers$House_norm_MP,
       col=centers$cluster, pch=as.numeric(centers$cluster),
       cex=5, lwd=3)
```
2. Based on the WSS plot (3.2) and the charts (4.1) choose one set of
   clusters that best describes the data.

#In the scatter plot, we can see that k=3 separates the data into three distinct clusters, with one cluster having low values for both Food and House spending, another cluster having moderate values for both, and a third cluster having high values for both.

#On the other hand, k=4 divides the data into four clusters, with two clusters having low values for both Food and House spending, another cluster having moderate values for both, and a fourth cluster having high values for Food and low values for House.

#If we want to focus on the overall spending pattern of the households and are interested in identifying households with high expenditure on both Food and House, then k=3 may be a better choice as it separates out a distinct cluster with high values for both. However, if we want to identify households with high Food expenditure but low House expenditure, then k=4 may be a better choice as it separates out a distinct cluster with these characteristics.

#Ultimately, I would prefer the cluster with k value 3.


3. Create summary tables for the segmentation/clustering scheme
   (selected in step 4.2).
```{r}

#scatter plot for k+1: 3+1=4
k=3

Clstr_MP <- kmeans(fcb_MP[,c(8,13)], iter.max=10, centers=k, nstart=10)

fcb_MP$cluster <- factor(Clstr_MP$cluster)

centers <- data.frame(cluster=factor(1:k), Clstr_MP$centers)

mss_MP[k-1] <- Clstr_MP$tot.withinss

summariseClst_MP <- aggregate(cbind(Food_MP, House_MP, Enter_MP, Work_MP, Edu_MP, Trans_MP, Oth_MP) ~ cluster, fcb_MP, FUN=mean)
summariseClst_MP

```

4. Create suitable descriptive names for each cluster.

#Cluster 1: BookWorms
#Cluster 2: Newyork
#Cluster 3: UrbanOwners

```{r}

cnames_MP <- c("BookWorms", "NewYork", "UrbanOwners")
summariseClst_MP$cluster <- cnames_MP
head(summariseClst_MP)

```

5. Suggest possible uses for this clustering scheme. 

#Pricing of insurance: Insurance providers can range their pricing strategies to each group in order to lower their exposure to underwriting risk and boost profitability.

#Pricing optimization: Based on each group's ability to pay, businesses can tailor their pricing strategies.

#Optimization of store layout: By grouping products that each cluster is more likely to purchase, businesses can use the clustering method to optimise the layout of their stores.

#Real estate: Real estate brokers can identify potential purchasers looking for upscale communities and residences using this clustering scheme and then assist in matching them with appropriate properties.
