---
title: "PROG8430-23W-Assign02"
author: "Maulik Patel"
date: "2023-02-10"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

#This sets the working directory
knitr::opts_knit$set(root.dir = setwd("C:/Users/mauli/OneDrive/Documents/BDSA/R/Assignment 2"))

```
This section is for the basic set up.
It clears all the plots, the console and the workspace.
It also sets the overall format for numbers.

```{r}
if(!is.null(dev.list())) dev.off()
#cat("\014") 
rm(list=ls())
options(scipen=12)
```
This section loads and attaches all the necessary packages.
```{r}
if(!require(pastecs)){install.packages("pastecs")}
library("pastecs")

if(!require(lattice)){install.packages("lattice")}
library("lattice")

if(!require(readxl)){install.packages("readxl")}
library("readxl")

if(!require(ggplot2)){install.packages("ggplot2")}
library("ggplot2")

if(!require(tinytex)){install.packages("tinytex")}
library("tinytex")

```

Data Transformation and Preparation

1.  Initial Transformation
    a.  Rename all variables with your initials appended (just as was done in assignment 1 and 2)
    
    ```{r}

#reading the file
df_MP <- read.table(file = "PROG8430-23W-Assign02.txt", header=TRUE, sep=",")

#transforming it into a Data frame
df_MP <- as.data.frame(df_MP)
head(df_MP)

#to check the datatype we used class() function
class(df_MP)

```

b.  Transform character variables to factor variables.

```{r}

#transforming character variables into factor variables using as.factor() function
df_MP$Manufacturer <- as.factor(df_MP$Manufacturer)
df_MP$Server <- as.factor(df_MP$Server)

#we will verify it using class() function here
class(df_MP$Manufacturer)
class(df_MP$Server)

```

```{r}

#Renaming all variables in table with my name initials
colnames(df_MP) <- c("Index_MP", "Manufacturer_MP", "Server_MP", "Conn_MP", 
                     "RC_MP", "ST_MP", "SMBR_MP", "SMBT_MP", "BR_MP", "BT_MP", 
                     "UC_MP", "FA_MP")

#after printing the data frame we can see that the names of the columns are updated with my initials
print(head(df_MP))


```

2.  reduce Dimensionality
    a.  Apply the Missing Value Filter to remove appropriate columns of data.
    
```{r}

#using summary() function we can have the overall summary of the data like min, 
#median, max and so on.
#using this we can clearly see that there are no NA's or Null values in the data.
summary(df_MP)


```
b.  Apply the Low Variance Filter to remove appropriate columns of data.

```{r}
#stat.desc() function gives us the descriptive statistical analysis of our data frame
#Cloumns Manufacturer and server are factor variables so it will return NA's 
#when we use stat.desc() function
stat.desc(df_MP)
```

```{r}

# Identifying the Low Variance in our data frame
#using stat.desc we can see that UC_MP has the lowest value of variance with 2.252003e-05, 
#so using the below command we will drop the column.
df_MP <- df_MP[-c(11)]
head(df_MP)

```
c.  Apply the High Correlation Filter to remove appropriate columns of data.

```{r}
#Here we created a new variable and dropped columns manufacturer and server 
#so we can calculate correlation of the data using cor() function.  
corr_MP <- df_MP[-c(2:3)]
cor(corr_MP)

```
```{r}
#Columns with high correlation needs to be dropped, if not it can create 
#complexities ahead.
#Here we dropped columns BR and BT because they were having high correlation 
#with the columns SMBT and SMBR
df_MP <- df_MP[-c(9:10)]
head(df_MP)

```

d.  Drop any variables that do not contribute any useful analytical information at all.


```{r}
#We have deleted the index column because it does not add any value to our data for analytics
df_MP <- df_MP[-c(1)]
head(df_MP)
```

3. Outliers
  a. Use an appropriate technique demonstrated in class to identify
     outliers.


```{r}
#appropriate technique demonstrated in class to identify outliers is using the boxplots
#Below we have plotted boxplots for each and every column to recognize outliers in the data

boxplot(df_MP$Conn_MP, horizontal=TRUE, pch=19, main='Conn_MP', outpch = 25, outbg = "tomato1")
boxplot(df_MP$RC_MP, horizontal=TRUE, pch=19, main='RC_MP', outpch = 25, outbg = "tomato1")
boxplot(df_MP$ST_MP, horizontal=TRUE, pch=19, main='ST_MP', outpch = 25, outbg = "tomato1")
boxplot(df_MP$SMBR_MP, horizontal=TRUE, pch=19, main='SMBR_MP', outpch = 25, outbg = "tomato1")
boxplot(df_MP$SMBT_MP, horizontal=TRUE, pch=19, main='SMBT_MP', outpch = 25, outbg = "tomato1")
boxplot(df_MP$FA_MP, horizontal=TRUE, pch=19, main='FA_MP', outpch = 25, outbg = "tomato1")


#For more detailed view of the plots we will adjust the 
#Inter Quartile Range of the plots using range() function

boxplot(df_MP$RC_MP, 
        horizontal=TRUE, 
        pch=19, 
        main='Adjusted IQR for RC_MP', 
        range=0.50, 
        outpch = 25, 
        outbg = "tomato1")

```

```{r}

#Density plots for variables

densityplot( ~ df_MP$Conn_MP, 
             pch=4, 
             main='Density plot for Connections Made', 
             xlab='Connections Made')

densityplot( ~ df_MP$RC_MP, 
             pch=4, 
             main='Density plot for Reconnections Made', 
             xlab='Reconnections Made')

densityplot( ~ df_MP$ST_MP, 
             pch=4, 
             main='Density plot forSession Timed Out', 
             xlab='Session Timed Out')

densityplot( ~ df_MP$SMBR_MP, 
             pch=4, 
             main='Density plot for Server Message Blocks Received', 
             xlab='Server Message Blocks Received')

densityplot( ~ df_MP$SMBT_MP, 
             pch=4, 
             main='Density plot for Server Message Blocks Transmitted', 
             xlab='Server Message Blocks Transmitted')

densityplot( ~ df_MP$FA_MP, 
             pch=4, 
             main='Density plot for Files Accessed', 
             xlab='Files Accessed')


```
```{r}
#In the below code we removed two outliers for the Reconnections Made variable
densityplot( ~ df_MP$RC_MP, 
             pch=4, 
             xlab="Reconnections Made", 
             main="Density plot for Reconnections Made")

#created a new variable rc_MP and stored the values less than 0 and then 
#we dropped the values less than 0 which were the outliers
rc_MP <- which(df_MP$RC_MP < 0)
df_MP <- df_MP[-c(rc_MP),]

#This is an updated density plot in which we have data without the outliers
densityplot( ~ df_MP$RC_MP, 
             pch=4, 
             xlab="Reconnections Made", 
             main="Outliers Removed from Reconnections Made")

```

```{r}
#removing outliers for variable SMBT
densityplot( ~ df_MP$SMBT_MP, 
             pch=4, 
             main='Density plot for SMBT', 
             xlab='Server Message Blocks Transmitted')

#used min for the minimum row value for the outlier
sc_MP <- which(df_MP$SMBT_MP == min(df_MP$SMBT_MP)) 
df_MP <- df_MP[-c(sc_MP),]

#removed outliers
densityplot( ~ df_MP$SMBT_MP, 
             pch=4, 
             main='Density plot for SMBT', 
             xlab='Server Message Blocks Transmitted')


```
```{r}

#Density plots for Files Accessed 
densityplot( ~ df_MP$FA_MP, 
             pch=4, 
             main='Density plot for FA', 
             xlab='Files Accessed')

#used min to replace the outlier with the mean value
df_MP$FA_MP <- ifelse(df_MP$FA_MP == min(df_MP$FA_MP),
                      mean(df_MP$FA_MP),df_MP$FA_MP)

#density plot for Files Accessed variable with outliers removed
densityplot( ~ df_MP$FA_MP, 
             pch=4, 
             main='Density plot for FA', 
             xlab='Files Accessed')
```

Organizing Data
  1. Scatter Plots
      a. Create a histogram for Server Message Blocks Received.
      
```{r}

#histogram for Server Message Blocks Received
hist(df_MP$SMBR_MP,
     main = "Histogram for SMBR",
     xlab = "Server Message Blocks Received",
     col= "steelblue2"
     )

```
b. Create a histogram for Server Message Blocks Transmitted.
```{r}

#histogram for Server Message Blocks Transmitted
hist(df_MP$SMBT_MP,
     main = "Scatter Plot of SMBT",
     xlab = "Server Message Blocks Transmitted",
     col= "olivedrab1",
     )

```
c. Create a scatter plot showing the relationship between SMBR
   and SMBT. (note: SMBR should be on the x-axis, SMBT should
   be the y-axis)


```{r}

#Scatter plot of SMBR vs SMBT
plot(df_MP$SMBR_MP, df_MP$SMBT_MP,
     main = "Scatter Plot of SMBR and SMBT",
     xlab = "Server Message Blocks Received",
     ylab = "Server Message Blocks Transmitted",
     pch = 20,
     col= "darkorchid1",
     )

```
d. What conclusions, if any, can you draw from the chart?

#Direction and form of the relationship of SMBR and SMBT can be seen from the scatterplot above.

#Direction is positive as the value of both x and y is moving away from zero and it is increasing.

#so a positive correlation between SMBR and SMBT is observed

#as the values of both x and y are positive we can see a linear form of relationship between SMBR and SMBT.

e. Calculate a correlation coefficient between these two
variables. Why did you choose the correlation coefficient you
did? What conclusion you draw from it?

```{r}

cor(df_MP$SMBR_MP, df_MP$SMBT_MP, method = c("pearson"))

#we used pearson method to calculate correlation coefficient between SMBR and SMBT

#Using correlation function we can find out the strength of the relationship of SMBR and SMBT
#there is a strong linear relationship between SMBR and SMBT

#Reason for choosing Pearson method is because our both the variables SMBR and SMBT 
#are normally distributed(bell-shaped) as we can observe it in the histogram above, 
#so pearson is the best fit for such type of variables where the data is normally 
#distributed and provides a more complete description of the correlation of 
#association of the two variables. 

#0.762 shows that there is a strong linear relationship between SMBR and SMBT

```

Inference 
  1. Normality 
    a. Create a QQ Normal plot of for Sessions Timed Out. 
     
```{r}

#Here qqnorm function creates a Normal Q-Q plot for Session Timed Out variable
qqnorm(df_MP$ST_MP, main = "Sessions Timed Out", pch=17, col='seagreen1')

#qqline function is used to draw a line through Q-Q plot to make the linear 
#relationship clearly visible
qqline(df_MP$ST_MP)


```

b. Conduct a statistical test for normality on Sessions Timed Out.

```{r}
#we have conducted shapiro-wilk normality test to check whether the variable 
#session timed out is normally distributed or not.
#used sample function because shapiro test can only take 3 to 5000 rows for the test
shapiro.test(sample(df_MP$ST_MP, 5000))

```

c. Is Sessions Times Out normally distributed? What led you to this conclusion?
#No, the session timed out is not normally distributed.
#The p-value is less than 0.05 which states that it is an alternative hypothesis.


2. Statistically Significant Differences
  a. Compare Sessions Times Out between the two major
     Manufacturers in your data set using a suitable hypothesis
     test.
     
```{r}

shapiro.test(sample(df_MP$ST_MP, 5000))
#Performing shapiro-wilk test to get the p-value so we can decide the type of hypothesis

wilcox.test(df_MP$ST_MP ~ df_MP$Manufacturer_MP, data = df_MP, exact = FALSE)
#We will use wilcoxon test for comparing because from the above shapiro test 
#we get the p-value=0.00000000000000022 which is less than 0.05 so we will assume 
#it is not normally distributed.

```
b. Explain why you chose the test you did.
#The data is not normally distributed so that is why we will apply wilcoxon test

c. Do you have strong evidence that Sessions Times Out are
   different between Manufacturers?
#The value of P=0.00000000000000022 here is the strong evidence that the data 
#is in alternative hypothesis so yes, we can conclude that the sessions times 
#out are different between manufacturers


3. Multiple Statistical Differences
  a. Determine if Files Accessed varies by Server using ANOVA (statistical) 
     and a sequence of boxplots (graphical).
```{r}
#boxplot for Files Accessed and Server variables
boxplot(FA_MP ~ Server_MP, data=df_MP,
        main="Files Accessed using different server",
        range=0)

#we have done ANOVA test which is done to compare multiple means

#Here we are using one-way ANOVA test because we only have two variables FA and Server

#one-way ANOVA
summary(aov(FA_MP ~ Server_MP, data=df_MP))
print(" ")

ano_MP <- aov(FA_MP~Server_MP, data=df_MP)
summary(ano_MP)

#so, files accessed using different servers varies from server to server
```
  b. Determine if Connections Made varies by Server using ANOVA 
     and a sequence of boxplots.
```{r}
#boxplot for Connection and Server variables
boxplot(Conn_MP ~ Server_MP, data=df_MP,
        main="Connection made with servers",
        range=0)

#one-way ANOVA test
summary(aov(Conn_MP ~ Server_MP, data=df_MP))
print(" ")
ano2_MP <- aov(Conn_MP~Server_MP, data=df_MP)
summary(ano2_MP)

#From the output we conclude that Connections made from all the servers are equal

```